---
title: "Math7340 HW10"
author: "Chengbo Gu"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1 (20 points) Preprocessing a data set
##### Install the "ArrayExpress" package from Bioconductor. Load the yeast microarray data using R commands: 
##### library(ArrayExpress) 
##### yeast.raw = ArrayExpress('E-MEXP-1551')

##### (a) Preprocess the raw data set into an expression data set using: the “mas” background correction method, the “quantiles” normalization method, “pmonly” pm correction method and “medianpolish” summary method. Give the R command here for doing this task.

```{r results='hide', message=FALSE, warning=FALSE}
library(ArrayExpress)
library(affy)
```
```{r results='hide', message=FALSE, warning=FALSE}
yeast.raw <-  ReadAffy(celfile.path= 'E:/yeast' )
eset <- expresso(yeast.raw, 
                 bgcorrect.method="mas", 
                 normalize.method="quantiles", 
                 pmcorrect.method="pmonly", 
                 summary.method="medianpolish")
```

##### (b) Print out the mean expression values for the first five genes across all samples.
```{r}
firstFive <- exprs(eset)[1:5,]
means <- apply(firstFive, 1, mean)
means
```

##### (c) How many genes and how many samples are in the preprocessed expression data set?
```{r}
dim(eset)
```
There are 10928 genes and 30 samples in the preprocessed expression data set.


### Problem 2 (30 points) Searching Annotations
##### (a) What is the annotation package for the yeast data set in question 1?
##### Install the annotation package from Bioconductor.
```{r}
anno <- annotation(yeast.raw)
anno
# db <- paste(anno, ".db", sep="")
# source("https://bioconductor.org/biocLite.R")
# biocLite(db)
```
The annotation package for the yeast data set is "yeast2".

##### (b) Search the 1769308_at gene GO numbers related to Molecular Function (MF). How many GO numbers do you get?

```{r results='hide', message=FALSE, warning=FALSE}
library(yeast2.db)
library(annotate)
```
```{r}
go1769308 <- get("1769308_at", env = yeast2GO)
gonr <- getOntology(go1769308, "MF")
gonr
length(gonr)
```
There are 7 GO numbers.


##### (c) Find the GO parents of the GO IDs in part (b). How many GO parents are there?

```{r results='hide', message=FALSE, warning=FALSE}
library(GO.db)
```
```{r}
gP <- getGOParents(gonr)
pa <- sapply(gP, function(x) x$Parents)
length(unique(pa))
```
There are 5 GO parents.

##### (d) Find the GO children of the GO IDs in part (b). How many GO children are there?
```{r}
gC <- getGOChildren(gonr)
ch <- sapply(gC, function(x) x$Children)
length(unique(unlist(ch)))
```
There are 434 GO children.

### Problem 3 (30 points) Gene filtering on B-cell ALL patients
##### We work with the patients in stages "B2","B3".

##### (a) We look for genes expressed differently in stages B2 and B3. Use genefilter to program the Wilcoxon test and the Welch t-test separately for each gene. For each test, we select the genes with p-value<0.001. To save computational time, we set exact=F in the Wilcoxon test function.

```{r results='hide', message=FALSE, warning=FALSE}
library(genefilter)
library(ALL)
```
```{r}
data(ALL)

patientB2 <- factor(ALL$BT %in% c("B2"))
patientB3 <- factor(ALL$BT %in% c("B3"))
wilcox <- function(x) ( wilcox.test (x[patientB2 == TRUE], x[patientB3 == TRUE], 
                                     paired = F, exact = F)$p.value < 0.001 )

selwilcox <- genefilter(exprs(ALL), filterfun(wilcox))

welch <- function(x) ( t.test(x[patientB2 == TRUE], x[patientB3 == TRUE], 
                              paired = F) $p.value < 0.001 )

selwelch <- genefilter(exprs(ALL), filterfun(welch))
```

##### (b) Compute a Venn diagram for the Wilcoxon test and the t-test, and plot it.
```{r results='hide', message=FALSE, warning=FALSE}
library(limma)
```
```{r}
x <- apply(cbind(selwilcox, selwelch), 2, as.integer)
vc <- vennCounts(x, include = "both")
vennDiagram(vc)
```


##### (c) How many pass the Wilcoxon filter? How many passes both filters?
```{r}
ALLwilcoxon <- ALL[selwilcox,]
nrow(exprs(ALLwilcoxon))

ALLboth <- ALL[selwilcox & selwelch,]
nrow(exprs(ALLboth))
```
We can infer from both the venn diagram and the commands above.

345 genes pass the Wilcoxon filter.

297 genes pass both filters.


##### (d) What is the annotation package for the ALL data set? Find the GO numbers for "oncogene".

```{r results='hide', message=FALSE, warning=FALSE}
library(annotate)
```

```{r}
annotation(ALL)
```
The annotation package for the ALL data set is "hgu95av2".

```{r results='hide', message=FALSE, warning=FALSE}
library(GO.db) 
library(hgu95av2.db)
```
```{r}
GOTerm2Tag <- function(term) { 
                GTL <- eapply(GOTERM, function(x) {grep(term, x@Term, value=TRUE)}) 
                Gl <- sapply(GTL, length)
                names(GTL[Gl>0])}
GOTerm2Tag("oncogene")
```
The GO number for "oncogene" is "GO:0090402".

##### (e) How many genes passing the filters in (a) are oncogenes?
```{r}
tran <- hgu95av2GO2ALLPROBES$"GO:0090402"
inboth <- tran %in% row.names(exprs(ALLboth)) 
ALLtran <- ALLboth[tran[inboth],]
dim(ALLtran)
```
There is no such gene.


### Problem 4 (20 points)
##### Stages of B-cell ALL in the ALL data. Use the limma package to answer the questions below.
##### (a) Select the persons with B-cell leukemia which are in stage B1, B2, and B3.
```{r results='hide', message=FALSE, warning=FALSE}
library(limma)
library(ALL)
```
```{r}
data(ALL)
allB <- ALL[, which(ALL$BT %in% c("B1", "B2", "B3"))]
```

##### (b) Use the linear model to test the hypothesis of all zero group means. Use “topTable()” to report the top five genes with nonzero means in B3 group.
```{r}
design.ma <- model.matrix(~0 + factor(allB$BT))
colnames(design.ma) <- c("B1", "B2", "B3")
fit <- lmFit(allB, design.ma)
fit <- eBayes(fit)
print( topTable(fit, coef=3, number=5, adjust.method="fdr"), digits=4)
```

##### (c) Use two contrasts to perform analysis of variance to test the null hypothesis of equal group means. Do this with a false discovery rate of 0.01. How many differentially expressed genes are found? Use “topTable()” to report the top five genes that express differently among the three groups.
```{r}
cont.ma <- makeContrasts(B1-B2,B2-B3, levels=factor(allB$BT))
fit1 <- contrasts.fit(fit, cont.ma)
fit1 <- eBayes(fit1)
dim(topTable(fit1, number=Inf, p.value=0.01, adjust.method="fdr"))
```
There are 314 genes that are expressed differentially.


```{r}
print( topTable(fit1, number=5, adjust.method="fdr"), digits=4)
```
